Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-06-21 2017-Xudong-SMT/submit-ICTAC2017.tex
Remove the CCS example. Restructure the S5.
Sketch to the head.
Section 5.1 -> Section 5.3.

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules, and a sketch of the algorithm
to compute such semantics, but also explains the enhancement we have
done to these definitions.
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

@@

Each state has its set of variables called state variables.
$State\ variables$ can only be modified by the assignment in the transitions targeting its owner state. 
A global state variable of a pLTS is a variable $x$ satisfying $\forall s\!\in\! S .\, x\!\in\! \vars(s)$.

@@

\TODO{I'm thinking about moving this second example to section 5...}

\subsection{CCS Parallel Example}

\begin{figure}[h]
  \centerline{\includegraphics[width=10cm]{XFIG/CCS1.eps}}
  \caption{pNet encodings for the CCS expression}  \label{schema:ccs-pnet}
\end{figure}

\QIN{
This CCS expression composing two operators, parallel and prefix. In Fig.~\ref{schema:ccs-pnet}, 
it is easy to recognize that the pNets $PN2$ and $PN3$ are similar to 
the sub-pNet in the LOTOS example in their structure and synchronization vectors. 
That is because they are the operators conduct a similar behavior from different algebras.
}

\QIN{
To present the parallel property, we have defined a set of synchronization vectors composing 
two subnets. We use the notation from the CCS to present the emit and receive action. 
The variables $v_1$, $v_2$ are the data transmitting between two components, 
showing its value-passing behavior.
}

@@

In our prototype we start directly with open pNets expressed
programmatically, using the pNet API (graphical editors may be added
later). From these, we compute
the open automaton of an open pNet. We explain here our main
implementation choices.
%:
%\begin{itemize}
%  \item Steps 2a-2b are merged, applying Tr2 from the
%premisses OTs from subnets to generate OTs without
%explicitely constructing the deduction tree.
%  \item We define a general naming schema for implementing the fresh 
%and clone functions with more structure than a simple global name
%generator. It ensures unicity, but builds names
%readable enough for debugging purpose.
%  \item Besides that, we also have designed a management of state variables and
%    assignments throughout the whole computation.
%%%   \item Satisfiability check is done only at
%%% top level of the open-transition construction. An alternative would be
%%% to submit the satisfiability check to the SMT solver at each level of
%%% the hierarchical OT construction, potentially reducing the overall number of
%%% combinations. But the submission to the SMT engine is costly, and more
%%% complexity analysis is required before deciding if this would be
%%% worthwhile. 
%\end{itemize}
%
%In the following sections, we detail:
%\begin{itemize}
%\item The structure of fresh variable names.
%\item The details of the implementation for step 2a-2b, refined into:
%  \subitem combining all possibilities of subnets activity/inactivity,
%  \subitem matching each combination with synchronization vectors.
%\item Management of local variables of pLTSs.
%\item Pruning unsatisfiable OTs using the Z3 SMT solver.
%\end{itemize}

\subsection{Combining and Matching using TR2}
\TODO{Xudong: move here a (new version) of the alogorithm loplevel sketch}
\QIN{
In previous work, we already have a sketch for computing the open automaton.
}

@@

@@

\QIN{
The final result open transitions represent all the
possible movements of the LOTOS example.
What can be observed is always the actions of the process $P$ from the beginning, until $P$ performs a $\delta(x)$ and the value of
$x$ is transmitted to the $acc(x);Q$ giving out a silent action. Such behavior is performed by the $ot_1$ and $ot_2$. 
Then it keeps presenting the behavior of $Q$. $ot_3$ has showed that. 
%The $ot_1$ transition means that the actions of the process $P$ could
%be transmitted out, until $P$ performs a $\delta(x)$ and the value of
%$x$ is transmitted, through transition
%$ot_2$. The other 3 open transitions shows the behaviors of the
%process $Q$ while the $ot_3$ and $ot_4$ won't be performed as the
%controller has already set the value $s_0$ to 1 in the previous
%action. 
}

@@

\subsubsection{Result of CCS example}
\QIN{
also can present all the possible behavior of that CCS formula with 16 open
transitions in Fig. ~\ref{schema:ccs-result}, here we omitted the details of open transitions. 
In this example, there are 32 more transitions detected unsatisfiable by Z3.
}

\QIN{
Here we only show one of the result transitions. It presents the value-passing between $a.P$ and $b.Q$ and they are synchronized then
give out an silent action $\tau$, process $P$ and $Q$ should not work in this case.
}

\begin{figure}[h]
  \centerline{\includegraphics[width=5cm]{XFIG/CCSresult}}

  $\begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{0 \xrightarrow{a}_{C_1} 1, 0 \xrightarrow{b}_{C_2} 1\}, ~~a=!ch_{11}(v_{11}) \!\land\! b=?ch_{11}(v_{11}) \!\land\! a\neq l \!\land\! b\neq r\!\land\! v_1=\tau}
    {\ostate{00} \xrightarrow{v_1} \ostate{11}}\\
  \end{array}$

\caption{The open automaton of the CCS parallel formula}  \label{schema:ccs-result}
\end{figure}

>>>>

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules
%, and a sketch of the algorithm
%to compute such semantics, but also explains the enhancement we have
%done to these definitions
.
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

@@

Each state has its set of variables called $State\ variables$, 
which can only be modified by the assignment in the transitions targeting its owner state. 
A global state variable of a pLTS is a variable $x$ satisfying $\forall s\!\in\! S .\, x\!\in\! \vars(s)$.

@@

\TODO{I'm thinking about moving this second example to section 5...}
\QIN{[This CCS example is removed.]}

%\subsection{CCS Parallel Example}
%
%\begin{figure}[h]
%  \centerline{\includegraphics[width=10cm]{XFIG/CCS1.eps}}
%  \caption{pNet encodings for the CCS expression}  \label{schema:ccs-pnet}
%\end{figure}
%
%\QIN{
%This CCS expression composing two operators, parallel and prefix. In Fig.~\ref{schema:ccs-pnet}, 
%it is easy to recognize that the pNets $PN2$ and $PN3$ are similar to 
%the sub-pNet in the LOTOS example in their structure and synchronization vectors. 
%That is because they are the operators conduct a similar behavior from different algebras.
%}
%
%\QIN{
%To present the parallel property, we have defined a set of synchronization vectors composing 
%two subnets. We use the notation from the CCS to present the emit and receive action. 
%The variables $v_1$, $v_2$ are the data transmitting between two components, 
%showing its value-passing behavior.
%}

@@

\QIN{
For the open pNets, in previous work, we already have a sketch for computing the open automaton.
}

@@

However, more details are needed when doing the implementation. 
In our prototype we start directly with open pNets expressed
programmatically, using the pNet API (graphical editors may be added
later). 
%From these, we compute
%the open automaton of an open pNet. We explain here our main
%implementation choices.
%:
%\begin{itemize}
%  \item Steps 2a-2b are merged, applying Tr2 from the
%premisses OTs from subnets to generate OTs without
%explicitely constructing the deduction tree.
%  \item We define a general naming schema for implementing the fresh 
%and clone functions with more structure than a simple global name
%generator. It ensures unicity, but builds names
%readable enough for debugging purpose.
%  \item Besides that, we also have designed a management of state variables and
%    assignments throughout the whole computation.
%%%   \item Satisfiability check is done only at
%%% top level of the open-transition construction. An alternative would be
%%% to submit the satisfiability check to the SMT solver at each level of
%%% the hierarchical OT construction, potentially reducing the overall number of
%%% combinations. But the submission to the SMT engine is costly, and more
%%% complexity analysis is required before deciding if this would be
%%% worthwhile. 
%\end{itemize}
%
%In the following sections, we detail:
%\begin{itemize}
%\item The structure of fresh variable names.
%\item The details of the implementation for step 2a-2b, refined into:
%  \subitem combining all possibilities of subnets activity/inactivity,
%  \subitem matching each combination with synchronization vectors.
%\item Management of local variables of pLTSs.
%\item Pruning unsatisfiable OTs using the Z3 SMT solver.
%\end{itemize}

\subsection{Combining and Matching using TR2}
\TODO{Xudong: move here a (new version) of the alogorithm loplevel sketch}
\QIN{[Move to the head of Section 5.]}

@@

\QIN{
The final result open transitions represent all the
possible movements of the LOTOS example.
What can be observed is always the actions of the process $P$ from the beginning, until $P$ performs a $\delta(x)$ and the value of
$x$ is transmitted to the $acc(x);Q$ giving out a silent action. Such behavior is performed by the $ot_1$ and $ot_2$. 
Then it keeps presenting the behavior of $Q$. This can also be seen in $ot_3$. 
%The $ot_1$ transition means that the actions of the process $P$ could
%be transmitted out, until $P$ performs a $\delta(x)$ and the value of
%$x$ is transmitted, through transition
%$ot_2$. The other 3 open transitions shows the behaviors of the
%process $Q$ while the $ot_3$ and $ot_4$ won't be performed as the
%controller has already set the value $s_0$ to 1 in the previous
%action. 
}

@@

%\subsubsection{Result of CCS example}
%\QIN{
%also can present all the possible behavior of that CCS formula with 16 open
%transitions in Fig. ~\ref{schema:ccs-result}, here we omitted the details of open transitions. 
%In this example, there are 32 more transitions detected unsatisfiable by Z3.
%}
%
%\QIN{
%Here we only show one of the result transitions. It presents the value-passing between $a.P$ and $b.Q$ and they are synchronized then
%give out an silent action $\tau$, process $P$ and $Q$ should not work in this case.
%}
%
%\begin{figure}[h]
%  \centerline{\includegraphics[width=5cm]{XFIG/CCSresult}}
%  
%  $\begin{array}{@{}r@{}l@{}}
%    ot_1  = &\openrule{\{0 \xrightarrow{a}_{C_1} 1, 0 \xrightarrow{b}_{C_2} 1\}, ~~a=!ch_{11}(v_{11}) \!\land\! b=?ch_{11}(v_{11}) \!\land\! a\neq l \!\land\! b\neq r\!\land\! v_1=\tau}
%    {\ostate{00} \xrightarrow{v_1} \ostate{11}}\\
%  \end{array}$
%  
%\caption{The open automaton of the CCS parallel formula}  \label{schema:ccs-result}
%\end{figure}
